Condition-Based Maintenance for Complex Systems\\ based on Current Component Status\\ and Bayesian Updating of Component Reliability

### 2016-10-20 cbm-rbd-paper.tex
Weibull component model and update with cens, two points by Simme Douwe

\def\taut{\tau(\vec{t})}
\def\ttau{\tilde{\tau}}
\def\ttaut{\ttau(\vec{t})}

@@

@@

***\cite{2016:walter-coolen}

@@

>>>>

\def\taut{\tau(\vec{t})}
\def\ttau{\tilde{\tau}}
\def\ttaut{\ttau(\vec{t})}
\def\tautk{\tau(\vec{t}_k)}

@@

***operation: recalculate $\tau^*$ at fixed dense grid of time points (quasi-continuous),
this makes sense even if nothing happens because the fact that components in the system still function
gives information on component distributions;
alternatively, do the recalculation every time when a component fails (failures contain the most information)

***model allows also to switch to corrective policy when $g(\tau)$ monotonely decreasing (then $\tau \to \infty$)

@@

We consider a system with components of $k=1,\ldots,K$ different types;
for each type $k$, there are $n_k$ exchangeable components in the system.

***example reliability block diagram

@@

***RUL model description below follows closely \cite{2016:walter-coolen},
which describes the same RUL model.
However, \cite{2016:walter-coolen} does not consider maintenance policies,
and focuses on a generalization of the RUL model using sets of priors for the component models
(effectively parametric P-boxes),
allowing for vague and partial prior specifications,
and sensitivity to prior-data conflict.

\subsection{Bayesian update of Weibull component models}
\label{sec:weibull}

Here we describe the model for the component lifetimes,
together with the Bayesian update procedure which allows to include
expert knowledge, component tests, and information from the currently monitored components in the system.

For each type $k$ component, we assume that the lifetime $T_{k,i}$ ($i=1,\ldots,n_k$, $k = 1, \ldots, K$)
is Weibull distributed with fixed shape parameter $\beta_k > 0$,
in short $T_{k,i} \mid \lambda_k \sim \wei(\beta_k,\lambda_k)$,
with density and cdf
\begin{align}
\label{eq:weibulldens}
f(t_{k,i} \mid \lambda_k) &= \frac{\beta_k}{\lambda_k} (t_{k,i})^{\beta_k-1} e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}}\,, \\
\label{eq:weibullcdf}
F(t_{k,i} \mid \lambda_k) &= 1 - e^{-\frac{(t_{k,i})^{\beta_k}}{\lambda_k}} = P(T_{k,i} \leq t_{k,i} \mid \lambda_k)\,,
\end{align}
where $\lambda_k > 0$ and $t > 0$.

The shape parameter $\beta_k$ determines whether the hazard rate is increasing ($\beta_k > 1$)
or decreasing ($\beta_k < 1$) over time.
For $\beta_k=1$, one obtains the Exponential distribution with constant hazard rate as a special case.
To keep things simple, we assume $\beta_k$ to be known.
(We plan to extend the model to learn also $\beta_k$ in a later step,
using, e.g., the discretized approach by \cite{1969:soland}.)
The scale parameter $\lambda_k$ can be interpreted through the relation
\begin{align}
\E[T_{k,i} \mid \lambda_k] &= \lambda_k^{1/\beta_k}\, \Gamma(1 + 1/\beta_k)\,.
\label{eq:lambdainterpret}
\end{align}
We will use this equation to convert expected lifetimes to $\lambda_k$ and vice versa.

With a Bayesian approach, one can express expert knowledge about the reliability of the components
by assigning a so-called prior distribution,
a distribution over the unknown parameter $\lambda_k$.
This prior distribution $f(\lambda_k)$ is then updated 
to the so-called posterior distribution $f(\lambda_k \mid \vec{t})$,
the distribution over $\lambda_k$ given the data $\vec{t}$,
via Bayes' rule
\begin{align*}
f(\lambda_k \mid \vec{t}) &\propto f(\vec{t}\mid\lambda_k) f(\lambda_k)\,.
\end{align*}
The posterior $f(\lambda_k \mid \vec{t})$ subsumes the information
from both expert knowledge and data,
and forms the basis for all inferences, like, e.g., predictions.

For the prior over $\lambda_k$,
a convenient choice is to use the inverse Gamma distribution,
which is commonly parametrized in terms of the parameters $a_k > 0$ and $b_k > 0$,
\begin{align}
f(\lambda_k\mid a_k,b_k) &= \frac{(b_k)^{a_k}}{\Gamma(a_k)} \lambda_k^{-a_k -1} e^{-\frac{b_k}{\lambda_k}}\,,
\label{eq:ig-def}
\end{align}
in short, $\lambda_k \mid a_k, b_k \sim \ig(a_k,b_k)$.
Here, we have added the prior parameters $a_k$ and $b_k$ in notation
to indicate that the prior over $\lambda$ depends on their values.

The inverse Gamma is convenient because it is a conjugate prior,
i.e., the posterior obtained by Bayes' rule is again inverse Gamma and thus easily tractable;
the prior parameters only need to be updated to obtain the posterior parameters,
so no numerical integation or simulation techniques are necessary.
Furthermore, this conjugacy holds also when right-censored observations are used for updating,
as indicated below.

In place of the usual parametrization in terms of $a_k$ and $b_k$,
we use the parameters $\nkz > 1$ and $\ykz > 0$
because they have a simple interpretation.
They are defined as
\begin{align}
\nkz &= a_k - 1 & &\text{ and}
&
\ykz &= b_k / \nkz,
\label{eq:abtony}
\end{align}
where $\ykz$ can be interpreted as the prior guess for the scale parameter $\lambda_k$,
as $\E[\lambda_k\mid\nkz,\ykz] = \ykz$.
Using \eqref{eq:lambdainterpret},
we can thus translate an expert's statement of expected component lifetime into a corresponding value for $\ykz$.
$\nkz$ can be seen as a prior strength or pseudocount,
this will become clear in the discussion of the update step below.

The parametrization in terms of $\nkz$ and $\ykz$ also clarifies the nature of the combination
of prior information and data through Bayes' rule.
In the conjugate setting,
applying Bayes' rule simply means that the prior parameters $\nkz$ and $\ykz$
are updated to posterior parameters, which we denote by $\nkn$ and $\ykn$, respectively.
Assume we observe $n_k = e_k + c_k$ component lifetimes,
where $e_k$ is the number of actual failure events,
and $c_k$ is the number of right-censored observations.
We denote the failure times by $t_{k,1}, \ldots, t_{k,e_k}$,
and the censoring times by $t^+_{k,1}, \ldots, t^+_{k,c_k}$,
and collect them in an observation vector $\vec{t}_k = (t_{k,1}, \ldots, t_{k,e_k}, t^+_{k,1}, \ldots, t^+_{k,c_k})$.
Then, the updated, posterior parameters are
\begin{align}
\nkn &= \nkz + e_k\,, 
&
\ykn &=  \frac{\nkz \ykz + \tautk}{\nkz + e_k}\,,
\label{eq:ig-update}
\end{align}
where $\tautk = \sum_{i=1}^{e_k} (t_{k,i})^\beta + \sum_{i=1}^{c_k} (t^+_{k,i})^\beta$.

From the simple update rule \eqref{eq:ig-update}, we see that
$\ykn$ is a weighted average of the prior parameter $\ykz$ and the maximum likelihood (ML) estimator $\tautk/e_k$,
with weights proportional to $\nkz$ and $e_k$, respectively.
$\nkz$ can thus be interpreted as a prior strength or pseudocount,
indicating how much our prior guess should weigh against the $e_k$ observed failure events.
Furthermore, $\V[\lambda\mid\nkz,\ykz] = (\ykz)^2 / (1 - 1/\nkz)$;
for fixed $\ykz$, a higher $\nkz$ indicates thus 
that more probability mass is concentrated around $\ykz$.

Using \eqref{eq:abtony} and \eqref{eq:ig-update}, the posterior distribution over $\lambda_k$ is
\begin{align}
\lambda_k \mid \nkz, \ykz, \vec{t}_k \sim \ig(\nkz + e_k + 1, \nkz \ykz + \tautk)\,.
\label{eq:ig-update-alpha}
\end{align}
%where we have added the prior parameters $\nkz$ and $\ykz$ in notation
%to emphasize that the posterior is a synthesis of prior information and data.
As this posterior \eqref{eq:ig-update-alpha} can be defined in terms of
the updated, posterior parameters $\nkn$ and $\ykn$,
we may also write
\begin{align*}
f(\lambda_k \mid \nkz, \ykz, \vec{t}_k) &= f(\lambda_k \mid \nkn, \ykn)\,.
\end{align*}
The iterative nature of Bayesian inference means that we can take the updated,
posterior parameters as new prior parameters and update them again using new data.
Indeed, we will use this property to use the extra information about components
that accumulates during operation of the system:
At any time $\tnow > 0$,
all non-failed components correspond to a right censored observation $\tpnow$;
any failed components instead contribute a fully observed, non-censored lifetime,
and both types of observations can be used in the update step \eqref{eq:ig-update}.
To keep notation simple,
we will below denote the parameters of the inverse Gamma distribution by $\nkz$ and $\ykz$,
regardless of them expressing expert information alone,
or stemming from the combination of expert information and test data.

\section{Reliability functions for complex systems using the survival signature}
\label{sec:sysrel}

The goal is to find, for a given monitored system, the system reliability curve $\Rsys(t)$ at current time $\tnow$,
and we will now show how the system reliability function can be efficiently obtained using the survival signature.
This system reliability gives us the RUL distribution on which we base our adaptive maintenance policy.

We consider systems of arbitrary reliability block diagram layout,
consisting of components of $K$ different types,
and there are $n_k$ components of type $k$ in the system.
This system is observed until time $\tnow$,
leading to censored observation of lifetimes of components within the system.

\section{Posterior predictive***}

For this, we need the posterior predictive distribution
of the number of components that function at times $t > \tnow$,
which we will derive .
